Estimating the association between Facebook adoption and well-being

### 2022-09-16 ms.Rmd
Use PNAS template

---
title: The association between Facebook adoption and well-being
shorttitle: Facebook and well-being
leftheader: Facebook and well-being
author: 
  - name: Matti Vuorre
    affiliation: 1
    address: Oxford Internet Institute, University of Oxford
    email: matti.vuorre@oii.ox.ac.uk
    role:
      - Design
      - Administration
      - Analysis
      - Visualization
      - Writing
  - name: Andrew K. Przybylski
    affiliation: 1
    corresponding: yes
    address: Oxford Internet Institute, University of Oxford
    email: andy.przybylski@oii.ox.ac.uk
    role:
      - Design
      - Administration
      - Funding
      - Resources
      - Supervision
      - Visualization
      - Writing

affiliation:
  - id: 1
    institution: Oxford Internet Institute, University of Oxford

authornote: |
  \noindent \textbf{This manuscript is not yet peer-reviewed.}

  This research was supported by the Huo Family Foundation and Economic and Social Research Council Grant ES/T008709/1. All code supporting this manuscript is available at X. This draft is available under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/) at X.

  The authors declare no conflicts of interest.

  The analytic code is available at (LINK). Researchers can contact Facebook (WHERE?) to access our code, the Facebook data, and a synthetic version of the Gallup data for reproducing and extending our analyses. 
  This study and methodology was reviewed and approved by the Oxford Internet Institute's Departmental Research Ethics Committee (SSH_OII_CIA_21_084).

abstract: | 
  Abstract.

keywords: well-being, social media
# wordcount: 
bibliography: "references.bib"
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
header-includes:
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: jou
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

```{r setup, include = FALSE}
# Packages
library(papaja)
library(scales)
library(knitr)
library(patchwork)
library(ggh4x)
library(brms)
library(ggstance)
library(bayestestR)
library(tidyverse)

# For ordering outcomes
oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences"
)

# Directories for files
dir.create("models", FALSE)
dir.create("out", FALSE)

# Document options
opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  include = FALSE,
  warning = FALSE,
  dpi = 300,
  fig.align = "center"
)

@@

# Plotting options
theme_set(
  theme_linedraw(
    base_size = 8
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(margin = margin(4, 4, 4, 4, "pt")),
      axis.text = element_text(size = rel(0.7)),
      axis.ticks = element_line(size = rel(0.4))
    )
)
W <- 24
```

```{r data-prepare}
# Load and merge processed data
d_fb <- read_csv("data/fb.csv")
d_gwp <- read_rds("data/gwp.rds") 
# d_gwp <- read_rds("data/gwp-SYNTHETIC.rds") # If real data not available
d_gwp <- d_gwp %>% 
  mutate(outcome = factor(outcome, levels = oo))
d_pop <- read_rds("data/population.rds")

# Convert fb metrics into country-year proportions
d_fb <- d_fb %>% 
  left_join(d_pop) %>% 
  mutate(
    dau = dau / pop,
    mau = mau / pop
  )

@@

# Model formula
model_dau <- bf(
  val | se(se, sigma = TRUE) ~
    (year + dau_cw) * age * sex + dau_cb +
    ((year + dau_cw) * age * sex | country),
  family = gaussian()
)
model_mau <- bf(
  val | se(se, sigma = TRUE) ~
    (year + mau_cw) * age * sex + mau_cb +
    ((year + mau_cw) * age * sex | country),
  family = gaussian()
)

fits <- dm %>% 
  select(outcome, year, country, sex, age, contains("_c"), val, se) %>% 
  group_by(outcome) %>% 
  nest() %>% 
  mutate(
    dau = map2(
      data, outcome,
      ~brm(
        model_dau,
        cores = 8, chains = 4, threads = 2,
        backend = "cmdstanr",
        adapt_delta = .95,
        iter = 3000,
        data = .x,
        file = str_glue("models/brm-dau-{.y}")
      )
    ),
    mau = map2(
      data, outcome,
      ~brm(
        model_mau,
        cores = 8, chains = 4, threads = 2,
        backend = "cmdstanr",
        adapt_delta = .95,
        iter = 3000,
        data = .x,
        file = str_glue("models/brm-mau-{.y}")
      )
    )
  )
```

# Results

```{r fig-data}
#| include: true
#| out.width: "\\textwidth"
#| fig.env: "figure*"
#| fig.height: 5.4
#| fig.cap: Time courses of key variables. A. Daily and monthly Facebook users in two age brackets for 72 countries from 2008 to 2019. B. Mean responses to three well-being scales in the Gallup World Poll from 72 countries from 2008 to 2019.

# Figure of all variables over time
p_a <- d_fb %>% 
  pivot_longer(c(dau, mau), names_to = "outcome", values_to = "val") %>% 
  mutate(outcome = if_else(outcome == "dau", "Daily FB users", "Monthly FB users")) %>% 
  mutate(panel = str_glue("{outcome} ({age})")) %>% 
  ggplot(aes(year, val, group = country)) +
  scale_y_continuous(
    "Percentage of population", 
    expand = expansion(.03),
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2), 
    minor_breaks = seq(2009, 2019, by = 2)
  ) +
  geom_line(size = .2, col = "gray30") +
  geom_smooth(group = 1, method = "gam", se = F) +
  facet_wrap("panel", nrow = 1) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  )

p_b <- d %>% 
  select(-c(mau, dau)) %>% 
  # Exclude rows where FB data existed but not GWP data
  drop_na(val) %>% 
  # Actual lowest age in GWP
  mutate(age = factor(age, labels = c("15-34", "35+"))) %>% 
  mutate(panel = str_glue("{outcome} ({age})")) %>% 
  ggplot(aes(year, val, group = country)) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(
    "Value", 
    expand = expansion(.01),
    breaks = extended_breaks()
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2),
    labels = ~str_replace(., "20", "'")
  ) +
  # Aggregate over sex
  geom_line(size = .2, col = "gray30") +
  geom_smooth(group = 1, method = "gam", se = F) +
  facet_grid(sex~panel, labeller = as_labeller(~str_replace(., "_", " "))) +
  theme(legend.position = "none")

(p_a / p_b) +
  plot_layout(heights = c(4, 6)) +
  plot_annotation(tag_levels = "A")
# ggsave("out/Fig-data.png", width = W, height = W*.66, units = "cm", dpi = 400)
```

```{r fig-forest-prepare}
# Calculate contrasts for each outcome
h_dau <- c(
  "Average" = "dau_cw = 0",
  "13-34_Female" = 
    "dau_cw + dau_cw:age1 * -1 + dau_cw:sex1 * -1 + dau_cw:age1:sex1 * 1 = 0",
  "13-34_Male" = 
    "dau_cw + dau_cw:age1 * -1 + dau_cw:sex1 * 1 + dau_cw:age1:sex1 * -1 = 0",
  "35+_Female" = 
    "dau_cw + dau_cw:age1 * 1 + dau_cw:sex1 * -1 + dau_cw:age1:sex1 * -1 = 0",
  "35+_Male" = 
    "dau_cw + dau_cw:age1 * 1 + dau_cw:sex1 * 1 + dau_cw:age1:sex1 * 1 = 0",
  "13-34" = "dau_cw + dau_cw:age1 * -1 = 0",
  "35+" = "dau_cw + dau_cw:age1 * 1 = 0",
  "Female" = "dau_cw + dau_cw:sex1 * -1 = 0",
  "Male" = "dau_cw + dau_cw:sex1 * 1 = 0",
  "Between-country" = "dau_cb = 0"
)
h_mau <- c(
  "Average" = "mau_cw = 0",
  "13-34_Female" = 
    "mau_cw + mau_cw:age1 * -1 + mau_cw:sex1 * -1 + mau_cw:age1:sex1 * 1 = 0",
  "13-34_Male" = 
    "mau_cw + mau_cw:age1 * -1 + mau_cw:sex1 * 1 + mau_cw:age1:sex1 * -1 = 0",
  "35+_Female" = 
    "mau_cw + mau_cw:age1 * 1 + mau_cw:sex1 * -1 + mau_cw:age1:sex1 * -1 = 0",
  "35+_Male" = 
    "mau_cw + mau_cw:age1 * 1 + mau_cw:sex1 * 1 + mau_cw:age1:sex1 * 1 = 0",
  "13-34" = "mau_cw + mau_cw:age1 * -1 = 0",
  "35+" = "mau_cw + mau_cw:age1 * 1 = 0",
  "Female" = "mau_cw + mau_cw:sex1 * -1 = 0",
  "Male" = "mau_cw + mau_cw:sex1 * 1 = 0",
  "Between-country" = "mau_cb = 0"
)
coefs <- fits %>% 
  ungroup() %>% 
  pivot_longer(c(dau, mau), names_to = "predictor", values_to = "fit") %>% 
  mutate(h = if_else(predictor == "dau", list(h_dau), list(h_mau))) %>% 
  mutate(
    world = map2(
      fit, h,
      ~hypothesis(.x, .y) %>% 
        .[["hypothesis"]] %>% 
        tibble()
    ),
    country = map2(
      fit, h,
      ~hypothesis(.x, .y, scope = "coef", group = "country") %>% 
        .[["hypothesis"]] %>% 
        tibble()
    )
  ) %>% 
  select(-fit, -h, -data)

coefs <- coefs %>% 
  pivot_longer(c(world, country)) %>% 
  unnest(value) %>% 
  mutate(country = if_else(name == "country", as.character(Group), "World")) %>% 
  mutate(country = fct_relevel(country, "World")) %>% 
  # Between country contrast not country-specific!
  filter(!(Hypothesis == "Between-country" & country != "World"))

# Both predictor and outcome are 0-1, so scale can be interpreted as % change as a function of % change
```

```{r fig-forest}
#| include: true

coefs %>%   
  filter(str_detect(Hypothesis, "_")) %>% 
  mutate(
    predictor = if_else(
      predictor == "dau", 
      "Daily FB users", 
      "Monthly FB users"
    )
  ) %>%
  separate(Hypothesis, c("Age", "Sex"), sep = "_") %>% 
  mutate(outcome = str_glue("{outcome} ({Age})")) %>% 
  mutate(
    # Order countries on increasing magnitude per panel
    country_order = interaction(predictor, outcome, country),
    country_order = as.numeric(fct_reorder(country_order, Estimate)),
    # But also ensure that average ("World") is in bottom
    country_order = if_else(
      country == "World", country_order-10000, country_order
    ),
    country_order = factor(country_order)
  ) %>% 
  ggplot(aes(Estimate, country_order, col = Sex, shape = Star)) +
  scale_y_discrete("Country", expand = expansion(.015)) +
  scale_x_continuous(
    "Estimated association"
  ) +
  scale_shape_manual(values = c(21, 19)) +
  scale_color_brewer("Sex", palette = "Set1") +
  scale_size_manual(values = c(0.75, 1.5)) +
  geom_vline(xintercept = 0, lty = 2, size = .25, col = "grey50") +
  geom_linerangeh(
    aes(xmin = CI.Lower, xmax = CI.Upper),
    size = .2, position = position_dodgev(.4)
  ) +
  geom_point(
    fill = "white", aes(size = name),
    position = position_dodgev(.6)
  ) +
  facet_grid2(
    predictor~outcome, independent = "all", scales = "free",
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  guides(shape = "none", size = "none") +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    legend.position = "bottom"
  )

ggsave("out/Fig-forest.png", width = W, height = W*.66, units = "cm", dpi = 400)
```

```{r tab-between}
#| include: true

# Between-country associations
coefs %>% 
  mutate(
    predictor = if_else(
      predictor == "dau", 
      "Daily FB users", 
      "Monthly FB users"
    )
  ) %>%
  mutate(outcome = str_replace(outcome, "_", " ")) %>% 
  filter(Hypothesis == "Between-country") %>% 
  select(
    Predictor = predictor, Outcome = outcome, Estimate, CI.Lower, CI.Upper
  ) %>% 
  arrange(Predictor, Outcome) %>% 
  apa_table()
```

```{r fig-ce-prepare}
# Create a predictor table with evenly spaced mau_cw values for each country
newx_country <- fits$data[[1]] %>% 
  group_by(country, age, sex, mau_cb) %>% 
  filter(mau_cw %in% c(min(mau_cw), max(mau_cw))) %>% 
  select(country, age, sex, mau_cb, mau_cw) %>% 
  # For above groupings, create a sequence of dau_cws from min to max
  group_modify(
    ~tibble(mau_cw = seq(min(.$mau_cw), max(.$mau_cw), length = 31))
    ) %>% 
  ungroup() %>% 
  mutate(year = 0, se = 0)

>>>>

---
title: Estimating the association between Facebook adoption and well-being

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
author:
  - name: Matti Vuorre
    affiliation: a
  - name: Andrew K. Przybylski
    affiliation: a,1
address:
  - code: a
    address: Oxford Internet Institute, University of Oxford, 1 St Giles, Oxford OX1 3JS, United Kingdom

corresponding_author:
  code: 1
  text: "To whom correspondence should be addressed. E-mail: andy.przybylski@oii.ox.ac.uk"

# For footer text
lead_author_surname: Vuorre

author_contributions: |
  The authors made the following contributions. Matti Vuorre: Design, Administration, Analysis, Visualization, Writing; Andrew K. Przybylski: Design, Administration, Funding, Resources, Super- vision, Visualization, Writing.

## Remove this if not required
conflict_of_interest: |
  The authors declare no conflicts of interest.

abstract: |
  Social media's potential effects on well-being have received considerable research interest, but much of past work is hampered by an exclusive focus on demographics in the Global North and inaccurate self-reports of social media engagement. We describe associations linking 72 countries' Facebook adoption to the well-being of 946,798 individuals from 2008 to 2019. We found no associations suggesting that the global penetration of social media has caused widespread psychological harm: Facebook adoption predicted life satisfaction and positive experiences positively, and negative experiences negatively, both between countries and within countries but over time. Facebook adoption was more positively associated with well-being for younger individuals, but country-specific results were mixed. In order to move beyond studying aggregates and to better understand social media's causal effects, we need more transparent collaborative research between independent scientists and the technology industry.

significance: |
  Billions of individuals across the globe use social media on a regular basis. The widespread adoption of social media---most prominently Facebook---has fundamentally shifted how humans conduct their lives. Consequently, researchers, healthcare professionals, and policy makers have repeatedly raised alarms that the spread of social media may be negatively affecting individuals, particularly young ones, mental health and well-being on a global scale. We examined the plausibility of this idea by associating 72 nations' yearly Facebook adoption rates to the well-being of 946,798 individuals from 2008 to 2019. Overall, we found that increased rates of daily and monthly active Facebook users predicted greater levels of well-being, when comparing countries to other countries, or countries to themselves but over time. These findings should guide debate surrounding social media towards more empirical foundations: The importance of this topic requires that independent scientists gain access to more detailed data from the technology industry in order to better determine how and why modern online platforms might be affecting their users.

# acknowledgements: |
#   Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.

keywords:
  - well-being
  - social media

## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle

bibliography: references.bib
csl: pnas.csl

## change to true to add optional line numbering
lineno: true

output: 
  rticles::pnas_article:
    keep_tex: false
---

```{r setup, include = FALSE}
# Packages
library(papaja)
library(scales)
library(knitr)
library(patchwork)
library(ggh4x)
library(brms)
library(kableExtra)
library(ggstance)
library(bayestestR)
library(tidyverse)

# Directories for files
dir.create("models", FALSE)

# Document options
opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  echo = FALSE,
  include = FALSE,
  warning = FALSE,
  dpi = 300,
  fig.align = "center"
)

@@

# Plotting options
theme_set(
  theme_linedraw(
    base_size = 8
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(margin = margin(4, 4, 4, 4, "pt")),
      axis.text = element_text(size = rel(0.7)),
      axis.ticks = element_line(size = rel(0.4))
    )
)
```

```{r data-prepare}
# Load and merge data
# Facebook -- this data only exists on FORT
d_fb <- read_csv("data/fb.csv")
# If you are not conducting these analyses on FORT, the above data is not 
# available. Uncomment below to use a synthetic dataset instead.
# d_fb <- read_csv("data/fb-SYNTHETIC.csv")  # If real FB data not available

# Gallup
d_gwp <- read_rds("data/gwp.rds") 
# d_gwp <- read_rds("data/gwp-SYNTHETIC.rds") # If real Gallup data not available

# Population estimates
d_pop <- read_rds("data/population.rds")

# Convert fb metrics into country-year proportions
d_fb <- d_fb %>% 
  left_join(d_pop) %>% 
  mutate(
    dau = dau / pop,
    mau = mau / pop
  ) %>% 
  select(-pop)

@@

# Model formula
model_dau <- bf(
  val | se(se, sigma = TRUE) ~
    (year + dau_cw) * age * sex + dau_cb +
    ((year + dau_cw) * age * sex | country),
  family = gaussian()
)
model_mau <- bf(
  val | se(se, sigma = TRUE) ~
    (year + mau_cw) * age * sex + mau_cb +
    ((year + mau_cw) * age * sex | country),
  family = gaussian()
)
fits <- dm %>% 
  select(outcome, year, country, sex, age, contains("_c"), val, se) %>% 
  group_by(outcome) %>% 
  nest() %>% 
  crossing(
    nesting(
      model = list(model_dau, model_mau),
      predictor = c("dau", "mau")
    )
  ) %>% 
  mutate(
    fit = pmap(
      list(model, data, outcome, predictor),
      ~brm(
        ..1,
        cores = 4, chains = 4,
        backend = "rstan",
        control = list(adapt_delta = .95),
        iter = 3000,
        data = ..2,
        file = str_glue("models/brm-{..3}-{..4}")
      )
    )
  )
```

```{r coefficients-calculate}
path <- "models/coefficients.rds"
if (!file.exists(path)) {
  # Calculate contrasts for each outcome
  h_dau <- c(
    "Average" = "dau_cw = 0",
    "13-34_Female" = 
      "dau_cw + dau_cw:age1 * -1 + dau_cw:sex1 * -1 + dau_cw:age1:sex1 * 1 = 0",
    "13-34_Male" = 
      "dau_cw + dau_cw:age1 * -1 + dau_cw:sex1 * 1 + dau_cw:age1:sex1 * -1 = 0",
    "35+_Female" = 
      "dau_cw + dau_cw:age1 * 1 + dau_cw:sex1 * -1 + dau_cw:age1:sex1 * -1 = 0",
    "35+_Male" = 
      "dau_cw + dau_cw:age1 * 1 + dau_cw:sex1 * 1 + dau_cw:age1:sex1 * 1 = 0",
    "13-34" = "dau_cw + dau_cw:age1 * -1 = 0",
    "35+" = "dau_cw + dau_cw:age1 * 1 = 0",
    "Female" = "dau_cw + dau_cw:sex1 * -1 = 0",
    "Male" = "dau_cw + dau_cw:sex1 * 1 = 0",
    "Between-country" = "dau_cb = 0",
    "X by age" = "dau_cw:age1 = 0",
    "X by sex" = "dau_cw:sex1 = 0",
    "X by age by sex" = "dau_cw:age1:sex1 = 0"
  )
  h_mau <- c(
    "Average" = "mau_cw = 0",
    "13-34_Female" = 
      "mau_cw + mau_cw:age1 * -1 + mau_cw:sex1 * -1 + mau_cw:age1:sex1 * 1 = 0",
    "13-34_Male" = 
      "mau_cw + mau_cw:age1 * -1 + mau_cw:sex1 * 1 + mau_cw:age1:sex1 * -1 = 0",
    "35+_Female" = 
      "mau_cw + mau_cw:age1 * 1 + mau_cw:sex1 * -1 + mau_cw:age1:sex1 * -1 = 0",
    "35+_Male" = 
      "mau_cw + mau_cw:age1 * 1 + mau_cw:sex1 * 1 + mau_cw:age1:sex1 * 1 = 0",
    "13-34" = "mau_cw + mau_cw:age1 * -1 = 0",
    "35+" = "mau_cw + mau_cw:age1 * 1 = 0",
    "Female" = "mau_cw + mau_cw:sex1 * -1 = 0",
    "Male" = "mau_cw + mau_cw:sex1 * 1 = 0",
    "Between-country" = "mau_cb = 0",
    "X by age" = "mau_cw:age1 = 0",
    "X by sex" = "mau_cw:sex1 = 0",
    "X by age by sex" = "mau_cw:age1:sex1 = 0"
  )
  coefs <- fits %>% 
    select(-data, -model) %>% 
    mutate(h = if_else(predictor == "dau", list(h_dau), list(h_mau))) %>% 
    mutate(
      world = map2(
        fit, h,
        ~hypothesis(.x, .y) %>% 
          .[["hypothesis"]] %>%
          tibble()
      ),
      country = map2(
        fit, h,
        ~hypothesis(.x, .y, scope = "coef", group = "country") %>%
          .[["hypothesis"]] %>%
          tibble()
      )
    ) %>% 
    select(-fit, -h)

  # Both predictor and outcome are 0-1, so scale can be interpreted as % change as a function of % change

  coefs <- coefs %>% 
    pivot_longer(c(world, country)) %>% 
    unnest(value) %>% 
    mutate(
      country = if_else(name == "country", as.character(Group), "World")
    ) %>% 
    # Between country contrast not country-specific!
    filter(!(Hypothesis == "Between-country" & country != "World")) %>% 
    # Neatly formatted result string
    mutate(
      across(c(Estimate, CI.Lower, CI.Upper), list("str" = ~number(., .01))),
      result = str_glue("{Estimate_str}% [{CI.Lower_str}, {CI.Upper_str}]")
    )

  saveRDS(coefs, path)

} else {coefs <- readRDS(path)}
```

The ways in which people use technology for most domains in life has changed dramatically since the mass introduction of the internet in the 1990s, and the subsequent technologies facilitated by it. Most prominently, the popularization of modern social media platforms in 2008 precipitated widespread changes to human activities via features such as marketplaces, personalized news feeds, photo sharing, live streaming, and other features that the "metaverse" promises to build on. The first social media with global adoption, MySpace (launched 2003) saw 115 million users in 2008---the year in which it was replaced as the leading platform by Facebook (2004). Today (2022-Q1), Facebook reports 2.94 billion monthly active users [@metaMetaReportsFirst2022], or about one third of the global population. Along with social media's global penetration, debate surrounding their potential effects on individual and collective well-being has intensified.

Although reports of negative psychological outcomes associated social media are common in academic writing [@krossFacebookUsePredicts2013] and popular media [@thompsonSocialMediaAttention2021], evidence for harms is, on balance, more speculative than conclusive [@bestOnlineCommunicationSocial2014; @dickson2018screen; @odgersAnnualResearchReview2020]. Recent results on the associations between social media use and well-being are mixed and depend on arbitrary analytic choices [@orbenSocialMediaEnduring2019]. Other studies have reported that there have been few if any changes in associations linking technology use to mental health in this period of social media's global adoption [@vuorreThereNoEvidence2021]. A general lack of validated measures, poorly specified causal models, and inadequate data have yielded a large number of low-quality studies [@parrySystematicReviewMetaanalysis2021; @sewallDoesObjectivelyMeasured2022]. Furthermore, because nearly all investigations have focused on samples from the Global North [@ghaiSocialMediaAdolescent2022], we have next to no idea of how the wide adoption of social media platforms relates to psychological well-being across the world.

To better understand the plausible range of associations, we linked data tracking Facebook's global adoption with three indicators of global well-being. We examined 72 countries' per capita daily active Facebook users (DAU) in two age brackets (13-34 and 35+ years) as predictors of life satisfaction (LS), negative (NE), and positive psychological experiences (PE) at the level of years spanning 2008 to 2019. The well-being data represented 946,798 individuals' responses from the nationally representative Gallup World Poll Survey [@gallupHowDoesGallup2014]. We joined these unique datasets to conduct a descriptive study to answer three basic yet important questions: First, to what extent did Facebook adoption predict well-being? Second, do these associations differ by age or sex. And finally, how might these associations have differed between countries?

```{r tab-world}
#| echo: false
#| include: true

# Average within- and between-country associations
tab_world <- coefs %>% 
  filter(
    name == "world",
    Hypothesis %in% c("Average", "Between-country")
  ) %>% 
  mutate(
    outcome = case_when(
      outcome == "Life_satisfaction" ~ "LS",
      outcome == "Negative_experiences" ~ "NE",
      outcome == "Positive_experiences" ~ "PE"
    )
  ) %>% 
  select(
    predictor, Outcome = outcome, Hypothesis, result
  ) %>% 
  pivot_wider(names_from = Hypothesis, values_from = result) %>% 
  rename(
    `Within country` = Average,
    `Between country` = `Between-country`
  ) %>% 
  select(predictor, Outcome, `Between country`, `Within country`) %>% 
  arrange(predictor, Outcome)

tab_world %>% 
  filter(predictor == "dau") %>% 
  select(-predictor) %>% 
  mutate(
    across(
      ends_with("country"), 
      ~if_else(str_starts(., "-"), ., paste0("\\phantom{-}", .))
    )
  ) %>% 
  apa_table(
    caption = "Average Facebook DAU and well-being associations",
    font_size = "small",
    note = "LS: Life satisfaction, NE: Negative experiences: PE: Positive experiences. Numbers indicate posterior means and 95\\%CIs.",
    span_text_columns = FALSE,
    escape = FALSE,
    placement = "htb",
    align = 'llll'
  )
```

```{r}
# How many people in GWP data
d_gwp %>% 
  group_by(outcome) %>% 
  summarise(n = sum(n)) %>% 
  pull(n) %>% max %>% number

# Percentages of countries where DAU was over 25% by year
tmp <- d_fb %>% 
  group_by(year, age) %>% 
  summarise(median_dau = percent(median(dau))) %>% 
  pivot_wider(names_from = age, values_from = median_dau)
tmp
```

Facebook adoption increased markedly from 2008 when the median per capita DAU across these 72 countries was `r tmp[1,2]` (ages 13-34) and `r tmp[1,3]`% (35+), to `r tmp[12,2]` (13-34) and `r tmp[12,3]` (35+; Figure 1A) in 2019. During this period Facebook adoption by younger individuals plateaued before saturation in many countries, but not for older individuals. At the same time, we did not observe correspondingly large and uniform changes across measures of well-being (Figure 1B) [@vuorreGlobalWellbeingMental2022].

```{r fig-data, fig.pos = "H"}
#| include: true
#| fig.height: 3.6
#| fig.width: 3.5
#| fig.cap: "A. 72 countries' daily active Facebook users in two age brackets from 2008 to 2019. (Percentages may exceed 100\\% due to inaccurate estimation of either DAU or population size.) B. The same 72 countries' mean responses to three well-being scales in the Gallup World Poll from 2008 to 2019, separated by age category and sex. LS: Life satisfaction, NE: Negative experiences: PE: Positive experiences."

# Figure of all variables over time
p_a <- d_fb %>% 
  ggplot(aes(year, dau, group = country)) +
  scale_y_continuous(
    "Percentage of population", 
    expand = expansion(.03),
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2), 
    labels = ~str_replace(., "20", "'"),
    expand = expansion(0.01)
  ) +
  geom_line(size = .125) +
  # geom_smooth(group = 1, method = "gam", se = F) +
  facet_wrap("age", nrow = 1) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  ) +
  theme(text = element_text(size = 5))

p_b <- d %>% 
  # Exclude rows where FB data existed but not GWP data
  drop_na(val) %>% 
  # Actual lowest age in GWP
  mutate(age = factor(age, labels = c("15-34", "35+"))) %>% 
  mutate(
    outcome = case_when(
      outcome == "Life_satisfaction" ~ "LS",
      outcome == "Negative_experiences" ~ "NE",
      outcome == "Positive_experiences" ~ "PE"
    )
  ) %>% 
  mutate(panel = str_glue("{age} ({sex})")) %>% 
  ggplot(aes(year, val, group = country)) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(
    "Value", 
    expand = expansion(.01),
    breaks = extended_breaks(),
    labels = ~percent(.)
  ) +
  scale_x_continuous(
    "Year", 
    breaks = seq(2008, 2018, by = 2),
    labels = ~str_replace(., "20", "'")
  ) +
  geom_line(size = .1) +
  facet_grid(
    outcome~panel, 
    scales = "free_y",
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  theme(legend.position = "none") +
  theme(text = element_text(size = 5))

(p_a / p_b) +
  plot_layout(heights = c(3, 7)) +
  plot_annotation(tag_levels = "A")
```

To answer our first question of how Facebook adoption relates to well-being, we specified a multilevel linear meta-regression model of each well-being outcome on time, sex, age, and DAU (both within- and between-country-centered, see SI). We included all interactions between sex, age, and time, and sex, age, and DAU (within-country-centered), and allowed all parameters to vary randomly across countries. We first examined whether the relative standing of countries on their average Facebook adoption predicted well-being (the model's between-country associations). We found that countries with greater average DAU had somewhat higher levels of life satisfaction and positive experiences, and lower levels of negative experiences than countries with lower DAU (Table 1). However, there are large and important differences between countries in factors that might underlie differences in both social media adoption and well-being, such as socioeconomic conditions. As such, while descriptively informative, these associations are likely to indicate between-country confounding factors.

```{r fig-forest, fig.pos="H"}
#| echo: false
#| include: true
#| fig.height: 4.2
#| fig.width: 3.5
#| fig.cap: "Forest plot of estimated associations between country-year level per-capita daily active Facebook users (DAU) and life satisfaction (LS), negative experiences (NE), and positive experiences (PE). Units indicate percentage change in outcome as function of percentage increase in within-country centered DAU. Filled points indicate estimates whose posterior 95\\%CI excluded zero. See the online supplementary materials for corresponding figure but with MAU as predictor."

tmp <- coefs %>%   
  filter(
    # name != "world",
    str_detect(Hypothesis, "_")
  ) %>% 
  separate(Hypothesis, c("Age", "Sex"), sep = "_") %>% 
  mutate(
    outcome = case_when(
      outcome == "Life_satisfaction" ~ "LS",
      outcome == "Negative_experiences" ~ "NE",
      outcome == "Positive_experiences" ~ "PE"
    )
  ) %>% 
  # mutate(outcome = str_glue("{outcome} ({Sex})")) %>% 
  filter(predictor == "dau")

tmp %>%
  ggplot() +
  aes(
    x = Estimate, 
    y = fct_rev(country), 
    col = Age, 
    fill = interaction(Age, Star, sep = " "), 
    shape = Sex
  ) +
  scale_y_discrete(
    "Country",
    expand = expansion(.015),
    labels = ~str_replace(., "World", "Average")
  ) +
  scale_x_continuous(
    "Association with daily active Facebook users (95%CI)",
    breaks = extended_breaks(7),
    expand = expansion(.1)
  ) +
  scale_shape_manual(
    values = c(21, 22)
  ) +
  scale_color_brewer(
    "Age", palette = "Set1", aesthetics = c("color")
  ) +
  scale_fill_manual(values = c("white", "white", "#E41A1C", "#377EB8")) +
  geom_vline(xintercept = 0, lty = 2, size = .25, col = "grey50") +
  geom_linerangeh(
    aes(xmin = CI.Lower, xmax = CI.Upper),
    size = .15, position = position_dodge2v(.7), 
    show.legend = FALSE
  ) +
  geom_point(
    size = .65, stroke = .45,
    position = position_dodge2v(.7)
  ) +
  facet_wrap(
    "outcome", scales = "free_x", nrow = 1,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  guides(
    fill = "none",
    color = guide_legend(override.aes = list(size = 1.5)),
    shape = guide_legend(override.aes = list(size = 1.5))
  ) +
  theme(
    text = element_text(size = 5),
    axis.text.y = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    panel.grid.major.y = element_line(
      size = .1, linetype = 2, color = "gray60"
    ),
    legend.position = "bottom", 
    legend.box.margin = margin(-15, 0, -10, 0, "pt"),
    legend.background = element_rect(fill = NA),
    legend.key = element_rect(fill = NA),
    legend.text = element_text(size = rel(0.9))
  )
```

Variations in such confounding factors are likely to be significantly smaller within countries over time. We therefore next focused on the model's within-country associations, which measure the extent to which Facebook adoption in a given country predicted well-being in that country adjusting for general linear temporal trends in the well-being outcome. On average across countries, age, and sex, we found that years with greater than the average country's average DAU were associated with greater levels of life satisfaction and positive experiences, and lower levels of negative experiences (Table 1). For the average country in our sample, Facebook adoption positively predicted well-being. While these associations indicate relations within countries and adjust for confounders that vary linearly with time, by including time as a predictor, they are still susceptible to confounders and do not indicate causal relations. Rather, this association describes the fact that, all else being equal, years with greater Facebook adoption tended to be those with greater levels of well-being.

```{r demographics}
coefs %>% 
  filter(
    str_detect(Hypothesis, " by "), 
    name == "world"
  ) %>% 
  select(
    Predictor = predictor, Outcome = outcome, Hypothesis, result
  ) %>% 
  pivot_wider(names_from = Outcome, values_from = result)
coefs %>% 
  filter(
    Hypothesis %in% c("13-34", "35+", "Female", "Male"),
    name == "world"
  ) %>% 
  select(
    Predictor = predictor, Outcome = outcome, Hypothesis, result
  ) %>% 
  pivot_wider(names_from = Outcome, values_from = result)

tab_demo <- coefs %>% 
  filter(
    str_detect(Hypothesis, "_"),
    name == "world"
  ) %>% 
  select(
    Predictor = predictor, Outcome = outcome, Hypothesis, result
  ) %>% 
  pivot_wider(names_from = Outcome, values_from = result)
tab_demo
```
